Accessing the COVID-19 Death Rate in the US: While It Varies Across Different Income Levels, the Impact of Political Preference is Unignorable

### 2023-11-01 outputs/paper/paper.qmd
Update the reference

To understand and explore the COVID death rate, I will use five data sets from three different sources. American Community Survey [citation] collects the socio-economic information for each county, such as the mean household income. The Johns Hopkins University's public data set [citation] presents data on COVID-19 cases and deaths by county. Regarding electoral data, Fox News have comprehensive records of the 2020 US Federal Election, providing the vote counts for each party across all counties. These data can help us to assess the inequality of mortality caused by COVID across different countries.

My analysis will be focused on the variations of death rates across different counties to see how the counties with different socio-economic factors will affect the COVID-19 deaths. Moreover, I will incorporate a political dimension, especifically the voting rate in support of the Republican party in each county, to see whether a higher proportion of Republican voters in a county correlates with the COVID-19 death rate. Through this paper, I aim to construct a model where the COVID-19 death rate is the dependent variable and includes only those predictors that are statistically significant. This study can help US citizens understand the factors affecting the death rate and inform their voting decisions in 2024.

@@

As introduced in the introduction, there are three primary sources of data in the data: the American Community Survey (ACS), The Center for Systems Science and Engineering (CSSE) at John Hopkins University [citation], and Fox News [citation]. Each source provides distinct data dimensions that contribute to predicting the COVID-19 death rate in each US county.

@@

Fox News has posted the 2020 Election data, which was collected by Tony McGovern [citation]. It contains the county-level election results during the 2020 Federal Election. For each county, it includes the number of votes for the demographic and republican party, as well as the total votes.

@@

The way to clean each data set is different. For all the ACS data, I downloaded them using ACS API [citation] and only extracted the necessary variables using the codebook published by ACS [citation]. Additionally, I standardized the county names and transformed state names into their abbreviated forms. Regarding the COVID data sourced from JHU, I only select the cases and death numbers of US counties. Lastly, for the election data, I not only extracted the county name and transformed the state name to its abbreviation form but also calculated the number of people voting for the Republican party per 1000 voters.

@@

>>>>

To understand and explore the COVID death rate, I will use five data sets from three different sources. American Community Survey [@acs] collects the socio-economic information for each county, such as the mean household income. The Center for Systems Science and Engineering (CSSE) at John Hopkins University [@jhu] presents data on COVID-19 cases and deaths by county. Regarding electoral data, Fox News have comprehensive records of the 2020 US Federal Election, providing the vote counts for each party across all counties. These data can help us to assess the inequality of mortality caused by COVID across different countries.

My analysis will be focused on the variations of death rates across different counties to see how the counties with different socio-economic factors will affect the COVID-19 deaths. Moreover, I will incorporate a political dimension, especially the voting rate in support of the Republican party in each county, to see whether a higher proportion of Republican voters in a county correlates with the COVID-19 death rate. Through this paper, I aim to construct a model where the COVID-19 death rate is the dependent variable and includes only those predictors that are statistically significant. This study can help US citizens understand the factors affecting the death rate and inform their voting decisions in 2024.

@@

As introduced in the introduction, there are three primary sources of data in the data: the American Community Survey (ACS) [@acs], The Center for Systems Science and Engineering (CSSE) at John Hopkins University [@jhu], and Fox News [@fox]. Each source provides distinct data dimensions that contribute to predicting the COVID-19 death rate in each US county.

@@

Fox News has posted the 2020 Election data, which was collected by Tony McGovern [@mcgovern]. It contains the county-level election results during the 2020 Federal Election. For each county, it includes the number of votes for the demographic and republican party, as well as the total votes.

@@

The way to clean each data set is different. For all the ACS data, I downloaded them using ACS API [@acs_api] and only extracted the necessary variables using the codebook published by ACS [@acs_codebook]. Additionally, I standardized the county names and transformed state names into their abbreviated forms. Regarding the COVID data sourced from JHU, I only select the cases and death numbers of US counties. Lastly, for the election data, I not only extracted the county name and transformed the state name to its abbreviation form but also calculated the number of people voting for the Republican party per 1000 voters.

@@

```{r}
#| include: false
library(ggplot2)
library(maps)
states <- map_data("state")
counties <- map_data("county")
state_map <- data.frame(state.name = tolower(state.name), state.abb = state.abb)
counties$region <- state_map$state.abb[match(counties$region, state_map$state.name)]
counties$subregion <- tools::toTitleCase(counties$subregion)

mer <- merge(counties, data, by.x=c("subregion", "region"), by.y=c("county", "state"), all.x = TRUE)
mer$death_rate[is.na(mer$death_rate)] <- 0

ggplot(data=mer, mapping=aes(x=long, y=lat, group=group, fill=death_rate)) + 
  geom_polygon(color="black") +
  scale_fill_gradient(low="white", high="blue", name="Income") + 
  ggtitle('U.S. Map with Income by County') + 
  theme_minimal() + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  coord_fixed(1.3)

ggplot(mer, aes(x=long, y=lat, group=group)) + 
  geom_polygon(color="black", fill="gray") +
  coord_map() +
  theme_minimal()

duplicated_rows <- data[duplicated(data[c("county", "state")]),]
if (nrow(duplicated_rows) > 0) {
  print("There are duplicated rows!")
  print(duplicated_rows)
}
```
